Values Encoded in Machine Learning Research

### 2025-02-10 latex/paper.tex
updated keywords

\subsection{Keyword based analysis}
To analyze the values encoded in these papers, we adopted the 74 keywords used by Birhane et al. (2023) in their study of influential ML papers. In their paper
they write that they use 67 values, but in their annotation template for manual annotations (annotations.tsv) they include 74 values. In their results folder they
just mention 62 values. In order to search for as many values as possible we decided to proceed with the 74 values from the annotation file in their data folder.  
These keywords include terms related to performance, generalization, novelty, societal impact, and ethical considerations. 
Unlike Birhane et al., who relied on manual annotation, we developed and deployed an automated keyword-scanning algorithm to systematically identify the presence of 
these keywords across the abstracts, introductions, and conclusions of the selected papers.
The Table\ref{tab:keyword_table} in the Appendix provides a comprehensive list of the values and their associated keywords used in our analysis.
For example we did not only query for the additonal keyword ``Novelty'' but also for ``novel'' as these two
words express the same value and because we use a keyword search, sentences with the word ``novel'' would not be counted towards the keyword ``novelty''.  
For each keyword occurrence, contextual information was extracted to ensure accurate interpretation. 
This semi-automated process balanced the scale of automated analysis with the depth of manual review.

\subsection{Quantitative Summary}
We measured the prevalence of each keyword in the dataset, calculating the relative frequency of each value.

@@

\begin{table}[h]
    \centering
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Keywords Birhane Paper} & \textbf{additional keywords for automated keyword search} \\
        \hline
        Novelty & novel \\
        Simplicity & simple \\
        Generalization & generalisable \\
        Flexibility/Extensibility & flexible, flexibility, extensibility, extensible \\
        Robustness & robust \\
        Realistic output & realistic \\
        Formal description/analysis & formal, mathematical \\
        Theoretical guarantees & guarantee \\
        Approximation & approximate \\
        Quantitative evidence (e.g. experiments) & experiment \\
        Qualitative evidence (e.g. examples) & example \\
        Scientific methodology & scientific \\
        Controllability (of model owner) & control \\
        Human-like mechanism & human \\
        Low cost & cheap, cost \\
        Large scale & scale \\
        Generality & general \\
        Principled & principles \\
        Exactness & exact \\
        Preciseness & precise \\
        Concreteness & correct \\
        Automatic & automated \\
        Efficiency & efficient \\
        Building on classic work & classic work \\
        Building in recent work & recent work \\
        Unifying ideas or integrating components & unifying \\
        Identifying limitations & limitations \\
        Critique & criticism \\
        Understanding (for researchers) & understanding \\
        Used in practice/Popular & practice, popular \\
        Reproducibility & reproduce \\
        Easy to implement & implement \\
        Requires few resources & resources \\
        Parallelizability / distributed & parallelizability, parallelization, distributed \\
        Facilitating use (e.g. sharing code) & sharing code \\
        Scales up & scale up \\
        Applies to real world & real world \\
        Learning from humans & humans, learning \\
        Practical & practice \\
        Useful & usefulness \\
        Interpretable (to users) & interpretable \\
        Transparent (to users) & transparent, transparency \\
        Privacy & privacy, private \\
        Fairness & fair \\
        Not socially biased & social bias, socially bias, social, society \\
        User influence & user \\
        Collective influence & collective \\
        Deferral to humans & deferral \\
        Critiqability & criticism \\
        Beneficence & beneficable \\
        Respect for Persons & respect \\
        Autonomy (power to decide) & autonomy, autonome \\
        Explicability & explicable \\
        Respect for Law and public interest & respect for law, respect for public interest \\
        Security & secure \\
        Easy to work with & easy \\
        Realistic world model & world model \\
        Fast & speed \\
        \hline
    \end{tabular}
    \caption{Values and their associated keywords used in the analysis.}
    \label{tab:keyword_table}
\end{table}

>>>>

\subsection{Keyword based analysis}
To analyze the values encoded in these papers, we adopted the 74 keywords used by Birhane et al. (2023) in their study of influential ML papers. In their paper
they write that they use 67 values, but in their annotation template for manual annotations (annotations.tsv) they include 74 values. In their results folder they
just mention 62 values. In order to search for as many values as possible we decided to proceed with the 74 values from the annotation file in their data folder.  
These keywords include terms related to performance, generalization, novelty, societal impact, and ethical considerations. 
Unlike Birhane et al., who relied on manual annotation, we developed and deployed an automated keyword-scanning algorithm to systematically identify the presence of 
these keywords across the abstracts, introductions, and conclusions of the selected papers.
The Table\ref{tab:keyword_table} in the Appendix provides a comprehensive list of the values and their associated keywords used in our analysis.
For example we did not only query for the additonal keyword ``Novelty'' but also for ``novel'' as these two
words express the same value and because we use a keyword search, sentences with the word ``novel'' would not be counted towards the keyword ``Novelty''.  
For each keyword occurrence, contextual information was extracted to ensure accurate interpretation. 
This semi-automated process balanced the scale of automated analysis with the depth of manual review. One major limitation to this keyword based approach is 
that for some keywords this works better than for others and these keywords might then be under- or overrepresented in the analysis. For example for the keyword ``Fairness'' is 
is really likeli that the keyword search will find the occurences if the paper adresses this topic because they will most liekli use the word in the context. 
As a counter example the keyword `Èasy to implement'' might not be found as easily because the word ``easy'' is used in many other contexts as well and therefore we can not query for it.
There are many different words in witch the researchers could express that their model is easy to implement and they are hard to capture all by a simple keyword search. In order to keep that limitations as simple as possible
we try to query for many similar keywords and hope that this will cover most of the values.
Even tough the keyword search is not perfect, it is a good way to get a first overview of the values encoded in the papers and to get a first impression of the values. To stay in the scope of this seminar paper 
we therefor decided to use this approach but we want to explicitly mention this limitation.

\subsection{Quantitative Summary}
We measured the prevalence of each keyword in the dataset, calculating the relative frequency of each value. In Figure 1, we present the distribution of the values across all 200 papers from both conferences. The top values 
are Performance, Quantitative evidence, Qualitative evidence and Generality. We specifically highlited the values that the Brihane et al. paper groups as ``User rights'' and ``Ethical Principles'' in the figure.

@@

%%%%% muss geupdated werden 
\begin{table}[h]
    \centering
    \begin{tabular}{|l|l|}
        \hline
        \textbf{Keywords Birhane Paper} & \textbf{additional keywords for automated keyword search} \\
        \hline
        Novelty & novel \\
        Simplicity & simple \\
        Generalization & generalisable \\
        Flexibility/Extensibility & flexible, flexibility, extensibility, extensible \\
        Robustness & robust \\
        Realistic output & realistic \\
        Formal description/analysis & formal, mathematical \\
        Theoretical guarantees & guarantee \\
        Approximation & approximate \\
        Quantitative evidence (e.g. experiments) & quantitative \\
        Qualitative evidence (e.g. examples) & qualitative \\
        Scientific methodology & scientific \\
        Controllability (of model owner) & control \\
        Human-like mechanism & human \\
        Low cost & cheap, cost \\
        Large scale & scale \\
        Generality & general \\
        Principled & principles \\
        Exactness & exact \\
        Preciseness & precise \\
        Concreteness & correct \\
        Automatic & automated \\
        Efficiency & efficient \\
        Building on classic work & classic work \\
        Building in recent work & recent work \\
        Unifying ideas or integrating components & unifying \\
        Identifying limitations & limitations \\
        Critique & criticism \\
        Understanding (for researchers) & understanding \\
        Used in practice/Popular & practice, popular \\
        Reproducibility & reproduce \\
        Easy to implement & implement \\
        Requires few resources & resources \\
        Parallelizability / distributed & parallelizability, parallelization, distributed \\
        Facilitating use (e.g. sharing code) & sharing code \\
        Scales up & scale up \\
        Applies to real world & real world \\
        Learning from humans & humans, learning \\
        Practical & practice \\
        Useful & usefulness \\
        Interpretable (to users) & interpretable \\
        Transparent (to users) & transparent, transparency \\
        Privacy & privacy, private \\
        Fairness & fair \\
        Not socially biased & social bias, socially bias, social, society \\
        User influence \\
        Collective influence & collective \\
        Deferral to humans & deferral \\
        Critiqability & criticism \\
        Beneficence & beneficable \\
        Respect for Persons & respect \\
        Autonomy (power to decide) & autonomy, autonome \\
        Explicability & explicable \\
        Respect for Law and public interest & respect for law, respect for public interest \\
        Security & secure \\
        Easy to work with & easy \\
        Realistic world model & world model \\
        Fast & speed \\
        \hline
    \end{tabular}
    \caption{Values and their associated keywords used in the analysis.}
    \label{tab:keyword_table}
\end{table}
